validateHOT - Validate your Holdout Task

### 2023-10-24 paper/paper.Rmd
updated paper - 231024

---
title: 'validateHOT - Validate your Holdout Task'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 1, 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 1
citation_author: Schramm & Lichters.
date: 22 October 2023
year: 2023
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
---

@@

validateHOT combines both validation and market simulation in one package. validateHOT's advantages are the following: a) it helps you to decide which is the best model to proceed by validating it, b) it runs relevant market simulations that help to find the right product combinations, and finally, c) it is open source tool for function that are usually implemented in Sawtooth Software and are ofen a black-box for researchers and practitioners.

@@

# Overview and features

```{r}
data("MaxDiff")
HOT <- createHOT(
  data = MaxDiff, # data frame
  id = 1, # index unique identifier
  none = 19, # index of none alternative
  prod = 7, # no of alternatives in HOT excluding none
  prod.levels = list(3, 10, 11, 15, 16, 17, 18), # index of alternatives
  method = "MaxDiff", # method applied
  choice = 20 # column index of choice alternative
)
```

@@

```{r}
turf(
  data = MaxDiff,
  opts = c(Option_01:Option_16),
  none = none,
  size = 3,
  approach = "thres"
) %>% 
  head(., n = 5) %>% 
  mutate_if(is.numeric, round, 2)
```

>>>>

---
title: 'validateHOT - Validate your Holdout Task'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 1, 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters.
date: 22 October 2023
year: 2023
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
---

\setlength{\headheight}{63.55022pt}
\addtolength{\topmargin}{-0.95425pt}
\newcommand{\colcod}[1]{\texttt{\color{purple}#1}}

@@

validateHOT combines both validation and market simulation in one package. validateHOT's advantages are the following: a) it helps you to decide which is the best model to proceed by validating it, b) it runs relevant market simulations that help to find the right product combinations, and finally, c) it is open source tool for functions that are usually implemented in Sawtooth Software and are often a black-box for researchers and practitioners.

@@

validateHOT is a practical tool for Sawtooth Software users in industry as well as academia. It provides an open source solution for a) validating a validation/holdout task and therefore ensuring that the model has predictive validity; b) running market simulations. Other packages, for example, Metrics [@Metrics] provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix (see \autoref{tab:table1}). However, to put the Sawtooth export into the right format, the user needs some data wrangling which can be a barrier. Other packages mainly focus on the analysis of conjoint analysis (e.g., ChoiceModelR [@ChoiceModelR], choicetools [@choicetools], logitR [@logitr], bayesm [@bayesm] etc.). To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations is still missing. The goal of validateHOT is still to fill this research gap. validateHOT creates market shares based on *share of preferences* or *first choice rule*, which can be helpful to report results for research articles according to open science standards. In addition, it also provides the function \colcod{turf}, which was offered formerly by the turfR package, which is no longer available on the Comprehensive R Archive Network (see [CRAN](https://cran.r-project.org/web/packages/turfR/index.html)). So, currently practitioners and academics mainly have to stick to paid solutions, for example, Sawtooth Software.

# Key functions

The main function that creates the validation/holdout task is the \texttt{\color{purple}createHOT} function, which creates the total utility of each alternative by applying the additive utility model [@rao2014, p. 82]. The other functions in the package mainly focus on four different components which are outlined in \autoref{tab:table1}.

| Validation metrics | Confusion matrix | Market Simulations | Rescaling scores |
|:----------------:|:----------------:|:----------------:|:----------------:|
|     hitrate()      |    accuracy()    |    freqassort()    |    att_imp()     |
|        kl()        |       f1()       |     marksim()      |  prob_scores()   |
|       mae()        |   precision()    |      reach()       |                  |
|      medae()       |     recall()     |       turf()       |                  |
|       mhp()        |  specificity()   |                    |                  |
|       rmse()       |                  |                    |                  |

: Overview of main four components of validateHOT and their corresponding functions \label{tab:table1}

# Typical workflow

After running the Hierarchical Bayes estimation in Sawtooth Software, the raw utility scores should be exported. Both the raw utilities and the actual choice in the validation/holdout task needs be read into *R* (please see the vignette for more advice). In the following, we provide the workflow for a MaxDiff study (the vignette also provides detailed examples for a CBC as well as an ACBC).

We create the validation/holdout task, which has a total of 7 items (\texttt{\color{purple}prod}) plus the no-buy alternative (\texttt{\color{purple}none}). Moreover, we also define the method, the actual choice in the validation/holdout task as well as the column indexes of the alternatives in the validation/holdout task. Finally, we define variables that we would like to keep in the data frame (\colcod{varskeep}). \texttt{\color{purple}createHOT} returns a data frame which we need for subsequent analyses.

```{r}
data("MaxDiff") # read in the data
HOT <- createHOT(
  data = MaxDiff, # data frame
  id = 1, # index unique identifier
  none = 19, # index of none alternative
  prod = 7, # no of alternatives in HOT excluding none
  prod.levels = list(3, 10, 11, 15, 16, 17, 18), # index of alternatives
  method = "MaxDiff", # method applied
  choice = 20, # column index of choice alternative
  varskeep = 21
)
```

Next, to get the relevant validation metrics that are often reported in conjoint studies, for example, hit rate [e.g., @ding2005], mean hit probability [mhp, @voleti2017], or mean absolute error [mae, @wlömert2014], we only need to provide the data, the alternatives in the validation/holdout task (\texttt{\color{purple}opts}), and the actual choice (\texttt{\color{purple}choice}). Everything can be implemented in the tidyverse [@tidyverse] logic.

\pagebreak

@@

Next, validateHOT also provides the five metrics for the confusion matrix. The underlying logic hereby is that the user has to provide a no-buy alternative (\texttt{\color{purple}none}). validateHOT calculates, for example, how often a buy or no-buy was correctly predicted, therefore, it is testing whether the model correctly predicts general demand (exemplary showed by applying the \texttt{\color{purple}accuracy} function and results split by \texttt{\color{purple}Group}).

```{r}
accuracy(
  data = HOT, # data frame
  group = Group, # optional grouping variable
  opts = c(Option_1:None), # column names of alternatives
  choice = choice, # column name of choice
  none = None # column name 
)
```

Finally, we show two functions for market simulations, namely \texttt{\color{purple}marksim} as well as \texttt{\color{purple}turf}. First, we calculate the market shares based on the multinomial logit model [@McFadden1974]. Besides the aggregated shares, \texttt{\color{purple}marksim} also provides standard errors and the 95th confidence interval.

```{r}
marksim(
  data = HOT,
  opts = c(Option_1:None),
  method = "shareofpref"
)
```

\texttt{\color{purple}turf}, a "product line extension model" [@miaoulis1990, p. 29] is a tool to find the perfect assortment that creates the highest reach and a powerful tool for MaxDiff studies [@chrzan2019, p. 108]. To optimize the search for the optimal bundle, we could include as well the arguments \texttt{\color{purple}fixed}, to define alternatives that have to be part of the assortment, as well as \texttt{\color{purple}prohib}, to define prohibitions of combinations of items that should not be part of the assortment (please see the vignette for more details and also to see how to apply \texttt{\color{purple}turf} with data obtained by a likert scale).

For the following example, let us assume that you conducted an anchored MaxDiff analysis with 10 items (\colcod{opts}) and you want to find the best assortment with a size of 3 (\colcod{size = 3}). As a threshold (\colcod{none}) you use the anchor (no-buy alternative). Finally, you apply the threshold approach, that is that each alternative with a utility above the threshold (in our case \colcod{none}) is marked as purchase option. We reduce the output to the Top 5 combos based on the reach (a \colcod{1} indicates present in this assortment while \colcod{0} means not present).

```{r}
turf(
  data = MaxDiff,
  opts = c(Option_01:Option_10),
  none = none,
  size = 3,
  approach = "thres"
) %>% 
  head(., n = 5) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  t() %>% 
  as.data.frame() %>% 
  slice(-1) %>% 
  rename_all(., ~ paste0("Combo ", c(1:5)))
```
